Automated Topic Naming

### 2012-01-20 ese/ese.tex
100% audio edit -- I switched West Indies mode at 270 words per minute -- pretty cool :O

For the third set of word-lists, \textsf{exp3}, we extended the word-lists from \textsf{exp2} using WordNet similarity matches. 
Similarity in WordNet means siblings in a hypernym tree. 
We do not include these words here for space
considerations~\footnote{For our data repository visit \url{http://softwareprocess.es/nomen/}}. 
Wordnet similarity is a very broad match. For example, the label \emph{maintainability} is associated with
words \emph{ease} and \emph{ownership}, and the word \emph{performance} has a `sense' that refers to musical performances, which is obviously unrelated to software development. In general, as we proceed from word-lists in \textsf{exp1} to that in \textsf{exp3}, our lists become more generic.

@@

% Did a little validation here
Table \ref{tbl:wordlist} shows how many topics were labelled for
MaxDB, MySQL and PostgreSQL.  Notice how PostgreSQL had far fewer
unlabelled topics, PostgreSQL also was under going far more
development as we evaluated 7744 PostgreSQL commits versus 8664 MaxDB
commits and 6223 MySQL commits. 
In terms of change-log comment words PostgreSQL had 164724, while
MaxDB had 68203, and MySQL had 101081. This implies there were
more PostgreSQL
terms for the topic analyzer to produce topics with.
And thus we suspect that PostgreSQL topics were
flooded with many terms, indicating that perhaps we needed more 20 topics per month
for PostgreSQL as this result indicates there is a lot of overlap in
the topics.

@@

%\DONE{this discussion is about which DB? where do the numbers come from?}
For \textsf{exp1} the labels with the most topics
were
%XXX Neil                    V changed from 748 is that right?
\emph{correctness} (182/305/640, which represent MySQL, MaxDB and PostgreSQL topic counts, respectively) and \emph{testability} (121/238/625). 
We did not see many results for \emph{usability} (4/0/138) or
\emph{accuracy} (3/0/27), which were infrequently matched. Note the significantly higher result for \emph{usability} in PostgreSQL---this suggests a difference in how project developers are discussing usability, at least with respect to our analysis.
We also looked for correlations between our labels: excluding double
matches (self-correlation), our highest co-occurring terms were
\emph{verifiability} or \emph{correctness}
with \emph{traceability}, and \emph{testability} with \emph{correctness}.% (76 and 62 matches, respectively).
% See the following sections for our error analysis.

@@

\subsubsection{Analysis of the Semi-unsupervised Labelling} %--- %Based
				%on the labels, and our manual topic
				%labelling, 
%\noindent \textbf{Analysis of the unsupervised labelling} --- %Based on the labels, and our manual topic labelling, 
%we compared the results of the unsupervised word matching approach. 
For each quality in the high-level ISO 9126 taxonomy (namely, \emph{Maintainability, Usability, Reliability, Efficiency, Portability, Functionality}) we assessed whether the semi-unsupervised labels for a topic matched the manual annotations we created for the validation corpus. 
Recall that the manual annotations were not used to
train the labelling process.
As described in Section \ref{sec:wordlist} we used both ROC and F-1 measures to evaluate the performance of the classification.
Figure \ref{fig:maxdb-unsup-results} shows our ROC results for
PostgreSQL, MaxDB and MySQL. We omit plots of \textsf{exp1} due to poor results. We describe F-1 results in the text below.

@@

% EXP3
For \textsf{exp3}, the average F-measure (macro-F1) for MaxDB was $0.26$ with a range $0.11$ to
$0.47$, and $0.36$ for MySQL with a range of $0.10$ to $0.65$. 
PostgreSQL had an average of $0.42$ with a range of $0.31$ to $0.54$.
MaxDB had an average precision and recall of $0.16$ and $0.67$
while MySQL had $0.3$ and $0.48$. PostgreSQL had average precision and recall of $0.27$ and $0.95$, respectively.

@@

\subsubsection{Analysis of the Supervised Labelling}
% \label{sec:suplabelling}
Because our data-set consisted of word counts we expected Bayesian techniques to perform well. Bayesian techniques are often used in spam filtering, which is similarly interested in word distributions. 
We tried other learners that WEKA~\cite{weka09} provides, including rule learners, decision tree learners, vector space learners, and support vector machines.  
Figure \ref{fig:best-learn-per-tag} shows the performance of the best
performing learner per label: 
% We considered the best learner for a label to be 
the learner that had the highest ROC value for that label. 
The best learner is important because one uses a single learner per
label. When applying our technique, for each NFR one should select the
best performing learner possible. Given the tractability of these techniques, a tool which applied all learners and presented the best result should be feasible. 
%Figure \ref{fig:best-learn-per-tag} uses the ZeroR learner as a baseline, a common practice in machine learning,
%as ZeroR naively chooses the largest category all of the time. 
%ZeroR occasionally outperforms other learners; for labels which are not as common, this is to be expected because any miscategorization will hurt accuracy. 
%This is why we use ROC values, instead of accuracy, as they can better represent performance on labels which are not applicable to the majority of samples.

@@

%For MySQL (Figure \ref{fig:mysql-timeline}), we similarly validated our NFR patterns with external mailing list data. 
The development period of MySQL we studied  (Figure \ref{fig:mysql-timeline}) saw the first releases to be licensed under the GPL. 
Version 3.23.31 (January, 2001) was the production release (non-beta), and the timberline view shows a flurry of topics labelled with \emph{functionality} and
\emph{maintainability}. 
After this point, this version enters the maintenance phase of its life-cycle. 
In May 2001, there is an increase in the number of topics labelled with \emph{portability}. 
This might be related to release 3.23.38, which focused on Windows compatibility. 
Similarly, in August, 2002, both \emph{functionality} and \emph{portability} are frequent, and mailing list data suggests this is related to the release
of version 3.23.52, a general bug fix with a focus on security (a component of the \emph{functionality} NFR in the ISO9126 model). 
After this point, efforts shift to the newer releases (4.0, 4.1, 5.0) and subsequently becomes more functionality oriented.

@@

\begin{figure}
  \centering
  \includegraphics[width=.9\textwidth]{figures/postgresql-author-cluster}
  \caption{PostgreSQL Author Clusters: Authors clustered by euclidean
    distance of their NFR contributions. The outer rectangles (blue)
    reflect the clustering of authors into 2 clusters; while the inner rectangles
    (purple) reflect clustering of authors into 6 clusters. The height is a
    measure of cluster similarity, Euclidean distance, where smaller
    values are more similar. Individuals that share
    branches are more similar to each other individuals who do
    not. The longer the path, in terms of height, between two
    individuals, the more different their NFR contributions are.}
\label{fig:authorcluster}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=0.9\textwidth]{figures/self-sample-Both}
  \caption{Measured IRR versus IRR of random labeling
    simulations. Perfect IRR is 1.0.
    The red line and points highlight the inter-rater reliability of
    NFR topics labels
    measured on PostgreSQL. 
    The box-plots are the distribution of 100,000 IRR ratings between
    random simulation, drawn from our distribution of ratings, and our
    labelling of NFRs.
    Note how for 4 of the 7 NFRs it is
    distinctly higher than the median of the simulations.
  }
  \label{fig:irr}
\end{figure}

@@

\paragraph{Construct validity} -- we used only commit messages rather than
mail or bug tracker messages. 
To extend further we would need matching repositories for each project.
%Each case study was BecauseSince we required a seamless record, it was not possible to obtain
%other data sources. 
Possibly they would have influenced our results, but there would be a degree of correlation between the corpora.
 It is possible for a given label to occur across the arbitrary 30-day boundary we set. We suspect but have not proved that this is insignificant. 
%Conversely, construct validity was addressed in our validation as we used the files, topics and revisions to annotate the commit messages. 
%We used mailing-lists to verify the purpose behind some events.
Our taxonomy for software NFRs is subject to dispute, but seems to be
generally accepted.  A future approach should consider a different taxonomy, such as one created by surveying developers on what ``types" of tasks they work on.
Finally, there are exogenous sources, such as
in-person discussions, which we did not access.  %We did not show these results to the stakeholders.

@@

We primarily relied on ROC (the area under the receiver operating
characteristic curve), as our measurement to compare the effectiveness
of semi-unsupervised and supervised learners. ROC is less biased than
F-Measure in cross-folds
validation~\cite{flach-icml03,Forman:2010:ACS:1882471.1882479} in the
case of class imbalance. We
feel that the fixed point of $0.5$ for random results allows us to
better describe the predictability achieved by some of these
techniques against random and ZeroR learners. 
We also include
F-Measure results to allow for the reader to validate against
F-Measure if they are more comfortable with it. Thus because of
F-Measure biased handling of class-imbalance and the inherent
class-imbalance that our data suffers from we chose ROC.

>>>>

For the third set of word-lists, \textsf{exp3}, we extended the word-lists from \textsf{exp2} using WordNet similarity matches. 
Similarity in WordNet means siblings in a hypernym tree. 
We do not include these words here for space
considerations~\footnote{For our word lists visit \url{http://softwareprocess.es/nomen/}}. 
Wordnet similarity is a very broad match. For example, the label \emph{maintainability} is associated with
words \emph{ease} and \emph{ownership}, and the word \emph{performance} has a `sense' that refers to musical performances, which is obviously unrelated to software development. In general, as we proceed from word-lists in \textsf{exp1} to that in \textsf{exp3}, our lists become more generic.

@@

% Did a little validation here
Table \ref{tbl:wordlist} shows how many topics were labelled for
MaxDB, MySQL and PostgreSQL.  Notice how PostgreSQL had far fewer
unlabelled topics, PostgreSQL also was under going far more
development as we evaluated 7744 PostgreSQL commits versus 8664 MaxDB
commits and 6223 MySQL commits. 
In terms of change-log comment words PostgreSQL had 164724, while
MaxDB had 68203, and MySQL had 101081. This implies there were
more PostgreSQL
terms for the topic analyzer to produce topics with.
And thus we suspect that PostgreSQL topics were
flooded with many terms, indicating that perhaps we needed more than 20 topics per month
for PostgreSQL, as this result indicates there is a lot of overlap in
the topics.

@@

%\DONE{this discussion is about which DB? where do the numbers come from?}
For \textsf{exp1} the labels with the most topics
were
%XXX Neil                    V changed from 748 is that right?
\emph{correctness} (182/305/640, which represent MySQL, MaxDB and PostgreSQL topic counts, respectively) and \emph{testability} (121/238/625). 
We did not see many results for \emph{usability} (4/0/138) or
\emph{accuracy} (3/0/27), which were infrequently matched. Note the significantly higher result for \emph{usability} in PostgreSQL---this suggests a difference in how project developers are discussing usability, at least with respect to our analysis.
We also looked for correlations between our labels: excluding double
matches (self-correlation), our highest co-occurring labels were
\emph{verifiability} or \emph{correctness}
with \emph{traceability}, and \emph{testability} with \emph{correctness}.% (76 and 62 matches, respectively).
% See the following sections for our error analysis.

@@

\subsubsection{Analysis of the Semi-unsupervised Labelling} %--- %Based
				%on the labels, and our manual topic
				%labelling, 
%\noindent \textbf{Analysis of the unsupervised labelling} --- %Based on the labels, and our manual topic labelling, 
%we compared the results of the unsupervised word matching approach. 
For each quality in the high-level ISO 9126 taxonomy (namely, \emph{Maintainability, Usability, Reliability, Efficiency, Portability, Functionality}) we assessed whether the semi-unsupervised labels for a topic matched the manual annotations we created for the validation corpus. 
Recall that the manual annotations were not used to
train the labelling process.
As described in Section \ref{sec:wordlist} we used both ROC and F-measure measures to evaluate the performance of the classification.
Figure \ref{fig:maxdb-unsup-results} shows our ROC results for
PostgreSQL, MaxDB and MySQL. We omit plots of \textsf{exp1} due to poor results. We describe F-measure results in the text below.

@@

% EXP3
For \textsf{exp3}, the average F-measure (macro-F1) for MaxDB was $0.26$ with a range $0.11$ to
$0.47$, and $0.36$ for MySQL with a range of $0.10$ to $0.65$. 
PostgreSQL had an average F-measure of $0.42$ with a range of $0.31$ to $0.54$.
MaxDB had an average precision and recall of $0.16$ and $0.67$
while MySQL had $0.3$ and $0.48$. PostgreSQL had average precision and recall of $0.27$ and $0.95$, respectively.

@@

\subsubsection{Analysis of the Supervised Labelling}
% \label{sec:suplabelling}
Because our data-set consisted of word counts we expected Bayesian techniques to perform well. Bayesian techniques are often used in spam filtering, which is similarly interested in word distributions. 
We tried other learners that WEKA~\cite{weka09} provides, including rule learners, decision tree learners, vector space learners, and support vector machines.  
Figure \ref{fig:best-learn-per-tag} shows the performance of the best
performing learner per label: 
% We considered the best learner for a label to be 
the learner that had the highest ROC value for that label. 
The best learner is important because one uses a single learner per
label. When applying our technique, per each NFR one should select the
best performing learner possible. Given the tractability of these techniques, a tool which applied all learners and presented the best result should be feasible. 
%Figure \ref{fig:best-learn-per-tag} uses the ZeroR learner as a baseline, a common practice in machine learning,
%as ZeroR naively chooses the largest category all of the time. 
%ZeroR occasionally outperforms other learners; for labels which are not as common, this is to be expected because any miscategorization will hurt accuracy. 
%This is why we use ROC values, instead of accuracy, as they can better represent performance on labels which are not applicable to the majority of samples.

@@

%For MySQL (Figure \ref{fig:mysql-timeline}), we similarly validated our NFR patterns with external mailing list data. 
The development period of MySQL we studied  (Figure \ref{fig:mysql-timeline}) saw the first releases to be licensed under the GPL. 
Version 3.23.31 (January, 2001) was the production release (non-beta), and the time-line view shows a flurry of topics labelled with \emph{functionality} and
\emph{maintainability}. 
After this point, this version enters the maintenance phase of its life-cycle. 
In May 2001, there is an increase in the number of topics labelled with \emph{portability}. 
This might be related to release 3.23.38, which focused on Windows compatibility. 
Similarly, in August, 2002, both \emph{functionality} and \emph{portability} are frequent, and mailing list data suggests this is related to the release
of version 3.23.52, a general bug fix with a focus on security (a component of the \emph{functionality} NFR in the ISO9126 model). 
After this point, efforts shift to the newer releases (4.0, 4.1, 5.0) and subsequently becomes more functionality oriented.

@@

\begin{figure}
  \centering
  \includegraphics[width=.9\textwidth]{figures/postgresql-author-cluster}
  \caption{PostgreSQL Author Clusters: Authors clustered by Euclidean
    distance of their NFR contributions. The outer rectangles (blue)
    reflect the clustering of authors into 2 clusters; while the inner rectangles
    (purple) reflect clustering of authors into 6 clusters. The height is a
    measure of cluster similarity, Euclidean distance, where smaller
    values are more similar. Individuals that share
    branches are more similar to each other individuals who do
    not. The longer the path, in terms of height, between two
    individuals, the more different their NFR contributions are.}
\label{fig:authorcluster}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=0.9\textwidth]{figures/self-sample-Both}
  \caption{Measured IRR versus IRR of random labeling
    simulations. Perfect IRR is 1.0.
    The red line and points highlight the measured inter-rater reliability of
    NFR topics labels between annotators of PostgreSQL.
    The box-plots are the distribution of 100,000 IRR ratings between
    random simulation, drawn from our distribution of ratings, and our
    labelling of NFRs.
    Note how for 4 of the 7 NFRs the measure IRR is
    distinctly higher than the median of the simulations.
  }
  \label{fig:irr}
\end{figure}

@@

\paragraph{Construct validity} -- we used only commit messages rather than
mail or bug tracker messages. 
To extend further we would need matching repositories for each project.
%Each case study was BecauseSince we required a seamless record, it was not possible to obtain
%other data sources. 
Possibly they would have influenced our results, but there would be a degree of correlation between the corpora.
 It is possible for a given label to occur across the arbitrary 30-day boundary we set. We suspect but have not proved that this is insignificant. 
%Conversely, construct validity was addressed in our validation as we used the files, topics and revisions to annotate the commit messages. 
%We used mailing-lists to verify the purpose behind some events.
Our taxonomy for software NFRs is subject to dispute, but seems to be
generally accepted.  A future approach should consider a different taxonomy, such as one created by surveying developers on what ``types'' of tasks they work on.
Finally, there are exogenous sources, such as
in-person discussions, which we did not access.  %We did not show these results to the stakeholders.

@@

We primarily relied on ROC (the area under the receiver operating
characteristic curve), as our measurement to compare the effectiveness
of semi-unsupervised and supervised learners. ROC is less biased than
F-Measure in cross-folds
validation~\cite{flach-icml03,Forman:2010:ACS:1882471.1882479} in the
case of class imbalance. We
feel that the fixed point of $0.5$ for random results allows us to
better describe the predictability achieved by some of these
techniques against random and ZeroR learners. 
We also include
F-Measure results to allow for the reader to validate against
F-Measure if they are more comfortable with it. Thus because of
F-Measure's biased handling of class-imbalance and the inherent
class-imbalance that our data suffers from we chose ROC.
